### Summary of UTR and CDS lengths

Median CDS and UTR lengths, considering only UTRs with support from three
or more reads.

```{r table_01b, results='asis'}
# Table 1: Gene Structure Summary
table_01b <- gene_structure_summary(utr5_stats_combined_filtered$length,
                                    as.numeric(genes$size),
                                    utr3_stats_combined_filtered$length)
xkable(table_01b, caption='**Table 1b** Gene structure summary table.')
```

### Comparison of UTR lengths across stages

Median UTR length for each development stage.

```{r lengths_by_developmental_stage, results='asis'}
# median 5' and 3'UTR lengths, by developmental stage
utr5_summary <- utr5_stats_filtered %>% 
	group_by(stage) %>% 
	summarise(median=median(length, na.rm=TRUE))

utr3_summary <- utr3_stats_filtered %>% 
	group_by(stage) %>% 
	summarise(median=median(length, na.rm=TRUE))

table_02 <- cbind(
	stage=stages,
	utr5=(utr5_summary %>% filter(stage != 'combined'))$median,
	utr3=(utr3_summary %>% filter(stage != 'combined'))$median
)
colnames(table_02) <- c('stage', "5'UTR", "3'UTR")

xkable(table_02, caption='**Table 2**: Stage-specific UTR lengths.')
```

### Average lengths

```{r figure_1_average_lengths, fig.cap=sprintf('**Figure 1** Characterization of %s gene structure', CONFIG$parasite), warning=FALSE}
# Average 5'UTR length
p1 <- ggplot(utr5_stats_combined_filtered, aes(x=length)) + 
        geom_histogram(aes(y=..ncount..), binwidth=20, fill=CONFIG$colors[1], colour=CONFIG$stroke, alpha=0.75) +
        xlim(0, 1500) +
        labs(x="5'UTR length (nt)", y="Transcript density")

# Average 3'UTR length
p2 <- ggplot(utr3_stats_combined_filtered, aes(x=length)) + 
        geom_histogram(aes(y=..ncount..), binwidth=20, fill=CONFIG$colors[1], colour=CONFIG$stroke, alpha=0.75) +
        xlim(0, 1500) +
        labs(x="3'UTR length (nt)", y="Transcript density")

# Average Polypyrimidine tract length
p3 <- ggplot(polypyrimidine_tracts, aes(x=length)) + 
        geom_histogram(aes(y=..ncount..), binwidth=10, fill=CONFIG$colors[1], colour=CONFIG$stroke, alpha=0.75) +
        labs(x="Polypyrimidine tract length (nt)", y="Transcript density")

# Average CDS length
cds_lengths <- as.data.frame(as.numeric(genes$size))
colnames(cds_lengths) <- 'length'

p4 <- ggplot(cds_lengths, aes(x=length)) + 
        geom_histogram(aes(y=..ncount..), binwidth=20, fill=CONFIG$colors[1], colour=CONFIG$stroke, alpha=0.75) +
        xlim(0, 1500) +
        labs(x="CDS length (nt)", y="Transcript density")

# Average distance between primary SL site and polypyrimidine tract
p5 <- ggplot(polypyrimidine_tracts, aes(x=sl_dist)) +
        geom_histogram(aes(y=..ncount..), binwidth=10, fill=CONFIG$colors[1], colour=CONFIG$stroke, alpha=0.75) +
        xlim(0, 250) +
        labs(x="Distance between primary trans-splicing site and polypyrimidine tract (nt)", y="Transcript density")

# Average distance between primary poly(A) site and polypyrimidine tract
p6 <- ggplot(polypyrimidine_tracts, aes(x=polya_dist)) +
        geom_histogram(aes(y=..ncount..), binwidth=10, fill=CONFIG$colors[1], colour=CONFIG$stroke, alpha=0.75) +
        xlim(0, 1500) +
        labs(x="Distance between primary polyadenylation site and polypyrimidine tract (nt)", y="Transcript density")

# Average intergenic length (distance between SL and Poly(A) sites)
p7 <- ggplot(intercds_regions %>% filter(stage == 'combined'), 
             aes(x=intergenic_length)) +
        geom_histogram(aes(y=..ncount..), binwidth=10, fill=CONFIG$colors[1], colour=CONFIG$stroke, alpha=0.75) +
        xlim(0, 1500) +
        labs(x="Intergenic length (nt)", y="Transcript density")

grid.arrange(p1, p5, p2, p6, p3, p7, p4, ncol=2)

# normalize plot y-axis heights
#get_ylim <- function (plt) {
#    ggplot_build(plt)$panel$ranges[[1]]$y.range[2]
#}
#plt_ylim <- max(sapply(list(p1, p2, p3), get_ylim))

#grid.arrange(p1 + ylim(0, plt_ylim), 
#             p2 + ylim(0, plt_ylim),
#             p3 + ylim(0, plt_ylim), ncol=2)

#plt_ylim <- max(sapply(list(p4, p5, p6), get_ylim))

#grid.arrange(p4 + ylim(0, plt_ylim), 
#             p5 + ylim(0, plt_ylim), 
#             p6 + ylim(0, plt_ylim), ncol=2)
```

### Stage-specific UTR lengths

#### 5'UTR

```{r figure_x_5utr_lengths, warning=FALSE}
# Stage-specific distributions
ggplot(utr5_stats_filtered %>% filter(stage !='combined'), aes(x=length, group=stage, fill=stage)) + 
       geom_histogram(binwidth=20, colour=CONFIG$stroke) +
       scale_fill_manual(values=CONFIG$colors[1:length(CONFIG$stages)]) +
       xlim(0, 1500) +
       labs(x="5'UTR length (nt)", y="Transcript density") +
       facet_grid(stage ~.)
```

#### 3'UTR

```{r figure_x_3utr_lengths, warning=FALSE}
# Stage-specific distributions
ggplot(utr3_stats_filtered %>% filter(stage !='combined'), aes(x=length, group=stage, fill=stage)) + 
       geom_histogram(binwidth=20, colour=CONFIG$stroke) +
       scale_fill_manual(values=CONFIG$colors[1:length(CONFIG$stages)]) +
       xlim(0, 1500) +
       labs(x="3'UTR length (nt)", y="Transcript density") +
       facet_grid(stage ~.)
```

### Functional enrichment of genes with short/long UTR's

To see if their are any specific functions enriched among the genes with the
shortest and longest UTR's, we will perform functional enrichment for the upper
and lower 25% of genes in each category.

```{r functional_enrichment_prep}
# Load GO annotations
go_terms <- load_go_terms(orgdb, keys(orgdb, keytype='GID'), keytype='GID')

# Gene / GO term mapping
gene_go_mapping <- as.data.frame(unique(
    go_terms %>% select(GID, GO, ONTOLOGY)
))
colnames(gene_go_mapping) <- c('gene', 'category', 'ontology')
```

#### Long 5'UTRs

```{r functional_enrichment_long_5utrs, results='asis'}
# list of all genes with confident UTR'5 lengths
all_genes_5utr <- utr5_stats_combined_filtered$gene

# Gene (CDS) lengths; used for goseq bias correction
genes_with_known_5utrs <- genes[genes$ID %in% all_genes_5utr]
gene_lengths_5utr <- width(genes_with_known_5utrs)
names(gene_lengths_5utr) <- genes_with_known_5utrs$ID

# number of genes in upper/lower quartiles
quartile_num_genes <- round(nrow(utr5_stats_combined_filtered) / 4)

# gene ids for genes in the upper quartile of 5'UTR lengths
utr5_upper_quartile <- (utr5_stats_combined_filtered %>% arrange(-length) %>% 
                        select(gene) %>% head(quartile_num_genes))$gene

# measure enrichment
enriched <- test_gene_enrichment(utr5_upper_quartile, all_genes_5utr,
                                 gene_go_mapping, gene_lengths_5utr)

cat('\n**over-represented functions:**\n\n')
xkable(enriched %>% filter(over_represented_pvalue_adj < 0.05))

cat('\n**under-represented functions:**\n\n')
xkable(enriched %>% filter(under_represented_pvalue_adj < 0.05))
```

#### Short 5'UTRs

```{r functional_enrichment_short_5utrs, results='asis'}
# gene ids for genes in the lower quartile of 5'UTR lengths
utr5_lower_quartile <- (utr5_stats_combined_filtered %>% arrange(-length) %>% 
                        select(gene) %>% tail(quartile_num_genes))$gene

# measure enrichment
enriched <- test_gene_enrichment(utr5_lower_quartile, all_genes_5utr,
                                 gene_go_mapping, gene_lengths_5utr)

cat('\n**over-represented functions:**\n\n')
xkable(enriched %>% filter(over_represented_pvalue_adj < 0.05))

cat('\n**under-represented functions:**\n\n')
xkable(enriched %>% filter(under_represented_pvalue_adj < 0.05))
```

#### Long 3'UTRs

```{r functional_enrichment_long_3utrs, results='asis'}
# list of all genes with confident 3'UTR lengths
all_genes_3utr <- utr3_stats_combined_filtered$gene

# Gene (CDS) lengths; used for goseq bias correction
genes_with_known_3utrs <- genes[genes$ID %in% all_genes_3utr]
gene_lengths_3utr <- width(genes_with_known_3utrs)
names(gene_lengths_3utr) <- genes_with_known_3utrs$ID

# number of genes in upper/lower quartiles
quartile_num_genes <- round(nrow(utr3_stats_combined_filtered) / 4)

# gene ids for genes in the upper quartile of 3'UTR lengths
utr3_upper_quartile <- (utr3_stats_combined_filtered %>% arrange(-length) %>% 
                        select(gene) %>% head(quartile_num_genes))$gene

# measure enrichment
enriched <- test_gene_enrichment(utr3_upper_quartile, all_genes_3utr,
                                 gene_go_mapping, gene_lengths_3utr)

cat('\n**over-represented functions:**\n\n')
xkable(enriched %>% filter(over_represented_pvalue_adj < 0.03))

cat('\n**under-represented functions:**\n\n')
xkable(enriched %>% filter(under_represented_pvalue_adj < 0.03))
```

#### Short 3'UTRs

```{r functional_enrichment_short_3utrs, results='asis'}
# gene ids for genes in the lower quartile of 3'UTR lengths
utr3_lower_quartile <- (utr3_stats_combined_filtered %>% arrange(-length) %>% 
                        select(gene) %>% tail(quartile_num_genes))$gene

# measure enrichment
enriched <- test_gene_enrichment(utr3_lower_quartile, all_genes_3utr,
                                 gene_go_mapping, gene_lengths_3utr)

cat('\n**over-represented functions:**\n\n')
xkable(enriched %>% filter(over_represented_pvalue_adj < 0.05))

cat('\n**under-represented functions:**\n\n')
xkable(enriched %>% filter(under_represented_pvalue_adj < 0.05))
```
