---
title: Trypanosome Gene Structure Analysis (T. cruzi v0.1)
author: Keith Hughitt
---

Introduction
============

In previous work, *L. major* and *T. cruzi* RNA-Seq data was used to determine
UTR boundaries and splice leader and polyadenylation acceptor 
sites ([1](https://github.com/elsayed-lab/utr_analysis)), and to look for
evidence of alternative trans-splicing and poly-adenylation, and compute basic
statistics relating to SL and Poly(A) site usage 
([2](http://www.umiacs.umd.edu/~keith/research/2015/110-utr-lengths/)) across
on a genome-wide scale.

Here, we expand on these analyses and attempt to map out the gene structure
for each of the parasites, and explore some of the basic properties and features 
of the trypanosome gene structure.

```{r include=FALSE}
library(knitr)
knit('settings/tcruzi_ystrain.Rmd', quiet=TRUE, output=tempfile())
```

```{r knitr_settings, include=FALSE}
opts_knit$set(progress=FALSE, verbose=TRUE)
opts_chunk$set(fig.width=1080/96,
               fig.height=1080/96,
               dpi=96,
			   echo=CONFIG$verbose)
options(digits=4)
options(stringsAsFactors=FALSE)
options(knitr.duplicate.label='allow')
```

```{r load_libraries, message=FALSE, warning=FALSE}
library(GenomicRanges)
library(Biostrings)
library(rtracklayer)
library(ggplot2)
library(reshape2)
library(dplyr)
library(readr)
library(MASS)
source('child/helper.R')
```

```{r include=FALSE}
knit('child/load_data.Rmd', quiet=TRUE, output=tempfile())
```

Results
=======

Overview
--------

```{r fig.height=250/96, fig.cap="**Figure 1**: Average 5'UTR, CDS, and 3'UTR lengths across all development stages."}
# Summary of gene structure
utr5_summary <- utr5_stats %>% 
	group_by(stage) %>% 
	summarise(median=median(length, na.rm=TRUE))

utr3_summary <- utr3_stats %>% 
	group_by(stage) %>% 
	summarise(median=median(length, na.rm=TRUE))

median_5utr <- (utr5_summary %>% filter(stage == 'combined'))$median
median_cds  <- median(as.numeric(genes$size))
median_3utr <- (utr3_summary %>% filter(stage == 'combined'))$median

plot_gene_structure(median_5utr, median_cds, median_3utr)
```

CDS/UTRs
--------

### Summary of UTR and CDS lengths

```{r gene_structure_summary_table, results='asis'}
# Table 1: Gene Structure Summary
summary_dat <- gene_structure_summary(utr5_stats_combined$length,
                                      as.numeric(genes$size),
                                      utr3_stats_combined$length)
kable(summary_dat, caption='**Table 1** Gene structure summary table.')
```

### Comparison of UTR lengths across stages

Median UTR length for each development stage.

```{r lengths_by_developmental_stage, results='asis'}
df <- cbind(
	stage=stages,
	utr5=(utr5_summary %>% filter(stage != 'combined'))$median,
	utr3=(utr3_summary %>% filter(stage != 'combined'))$median
)
colnames(df) <- c('stage', "5'UTR", "3'UTR")

kable(df, caption='Table 2**: Stage-specific UTR lengths.')
```

```{r child='child/length_correlations.Rmd'}
```

### RNA-Seq coverage support for UTR boundaries

```{r coverage_mapped_read_densities, fig.show='hold', out.width='450px', fig.cap="**Figure 3**: Distribution of number of SL- and Poly(A)-containing reads."}
# number of reads containing SL- or Poly(A) sequence
hist(log(utr5_stats_combined$num_reads + 1), main='# SL-containing reads')
hist(log(utr3_stats_combined$num_reads + 1), main='# Poly(A)-containing reads')
```

```{r coverage_summary_table, results='asis'}
total_genes <- nrow(utr5_stats_combined)
nvals <- c(1, 2, 3, 4, 5, 10, 100, 1000) 

# % genes with N or more reads mapped to feature
sl_pcts <- c()
polya_pcts <- c()

for (nval in nvals) {
	sl_pcts <- append(sl_pcts, sum(utr5_stats_combined$num_reads >= nval) / total_genes)
}

for (nval in nvals) {
	polya_pcts <- append(polya_pcts, sum(utr3_stats_combined$num_reads >= nval) / total_genes)
}

sl_pcts    <- round(sl_pcts, 4) * 100
polya_pcts <- round(polya_pcts, 4) * 100

# summary data frame
df <- cbind(N=nvals, SL=sl_pcts, PolyA=polya_pcts)

kable(df, caption='**Table 4**: Percent genes containing N or more reads mapping to the specified feature.')
```

