---
title: Trypanosome Gene Structure Analysis (T. cruzi v0.3)
author: Keith Hughitt
output:
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: true
---

Introduction
============

In previous work, *L. major* and *T. cruzi* RNA-Seq data was used to determine
UTR boundaries and splice leader and polyadenylation acceptor 
sites ([1](https://github.com/elsayed-lab/utr_analysis)), and to look for
evidence of alternative trans-splicing and poly-adenylation, and compute basic
statistics relating to SL and Poly(A) site usage 
([2](http://www.umiacs.umd.edu/~keith/research/2015/110-utr-lengths/)) across
on a genome-wide scale.

Here, we expand on these analyses and attempt to map out the gene structure
for each of the parasites, and explore some of the basic properties and features 
of the trypanosome gene structure.

```{r include=FALSE}
library(knitr)
knit('settings/tcruzi_ystrain.Rmd', quiet=TRUE, output=tempfile())
```

```{r knitr_settings, include=FALSE}
opts_knit$set(progress=FALSE, verbose=TRUE)
opts_chunk$set(fig.width=1080/96,
               fig.height=1080/96,
               dpi=96,
			   echo=CONFIG$verbose)
options(digits=4)
options(stringsAsFactors=FALSE)
options(knitr.duplicate.label='allow')

# If rmarkdown.pandoc.to not specified, have it default to 'latex' output
if (is.null(opts_knit$get("rmarkdown.pandoc.to"))) {
    opts_knit$set(rmarkdown.pandoc.to='latex')
}
```

```{r load_libraries, message=FALSE, warning=FALSE}
library('GenomicRanges')
library('Biostrings')
library('rtracklayer')
library('ggplot2')
library('reshape2')
library('dplyr')
library('readr')
library('MASS')
source('child/helper.R')
```

```{r include=FALSE}
knit('child/load_data.Rmd', quiet=TRUE, output=tempfile())
```

Results
=======

Overview
--------

```{r child='child/gene_structure_overview.Rmd'}
```

CDS/UTRs
--------

```{r child='child/length_summary.Rmd'}
```

```{r child='child/length_correlations.Rmd'}
```

```{r child='child/rnaseq_coverage.Rmd'}
```

Intergenic Regions
------------------

- **Note**: temporarily skipping to speed up development; will re-enable
     later.
<!--
``{r child='child/intergenic_regions.Rmd'}
``
-->

Mapping statistics
------------------

```{r child='child/mapping_statistics.Rmd'}
```

System Info
===========

```{r}
sessionInfo()
```

