## Average gene RPKM expression by chromosome

```{r data_prep}
# Create RPKM-normalized version of counts
gene_lengths <- data.frame(gene=genes$ID, length=as.numeric(genes$size))

# remove duplicates and entries missing from main count table
gene_lengths <- gene_lengths[gene_lengths$gene %in% rownames(count_table),]
gene_lengths <- gene_lengths[!duplicated(gene_lengths$gene),]

# Exclude counts for SLRNAs, etc. not included in gene_lengths
count_table <- count_table[rownames(count_table) %in% gene_lengths$gene,]

# Create a vector of gene lengths
length_vector <- gene_lengths[match(row.names(count_table), gene_lengths$gene), 'length']

rpkm_counts <- rpkm(count_table, gene.length=length_vector)
```

```{r figure_x_average_chr_gene_rpkm_expr}
# Get total expression for each gene, across all samples
rpkm_totals <- rownames_to_column(as.data.frame(rowSums(rpkm_counts)))
colnames(rpkm_totals) <- c('gene', 'expr')

# Add chromosome information
rpkm_totals <- merge(rpkm_totals, chr_mapping, by='gene') 

# Compute average total RPKM for each chromosome
chr_expr_averages <- rpkm_totals %>% 
    group_by(chr) %>%
    summarize(average_expr=median(expr),
              total_expr=sum(expr))

ggplot(chr_expr_averages, aes(x=chr, y=average_expr)) + 
    geom_bar(stat='identity', color=CONFIG$stroke, fill=CONFIG$colors[1]) +
    scale_x_discrete(limits=CONFIG$chr_order) +
    theme(axis.text.x=element_text(angle=90)) +
    xlab("Chromosome") +
    ylab("Median Gene Expression (RPKM)") +
    ggtitle("Average RPKM expression by chromosome")
```

## Average RPKM expression by chromosome at the nucleotide level

```{r figure_x_average_chr_rpkm_expr_by_nt}
chr_cds_total_lengths <- merge(gene_lengths, chr_mapping, by='gene') %>%
    group_by(chr) %>%
    summarise(total_length=sum(length))

chr_stats <- merge(chr_cds_total_lengths, chr_expr_averages, by='chr') %>%
    mutate(expr_per_nt=total_expr/total_length)

ggplot(chr_stats, aes(x=chr, y=expr_per_nt)) +
    geom_bar(stat='identity', color=CONFIG$stroke, fill=CONFIG$colors[1]) +
    scale_x_discrete(limits=CONFIG$chr_order) +
    theme(axis.text.x=element_text(angle=90)) +
    xlab("Chromosome") +
    ylab("Median RPKM expression per nucleotide") +
    ggtitle("Average RPKM expression per nt for each chromosome")
```

## Average RPKM expression by chromosome (extracellular vs. intracellular)

```{r figure_x_average_chr_gene_rpkm_expr_stage_specific}
# Create a long version of RPKM counts
rpkm_counts_long <- melt(rpkm_counts)
colnames(rpkm_counts_long) <- c('gene', 'sample_id', 'expr')

# Add stage information
rpkm_counts_long <- merge(rpkm_counts_long, CONFIG$samples %>%
                          select(sample_id, condition, stage),
                          by='sample_id')

# Add chromosome information
rpkm_counts_long <- merge(rpkm_counts_long, chr_mapping, by='gene') 

# Compute average stage-specific median RPKM for each chromosome 
# (intracellular vs. extracellular)
chr_expr_averages <- rpkm_counts_long %>% 
    group_by(chr, stage) %>%
    summarize(average_expr=median(expr),
              total_expr=sum(expr))

ggplot(chr_expr_averages, aes(x=stage, y=average_expr, fill=stage)) + 
    geom_bar(stat='identity', color=CONFIG$stroke) +
    scale_fill_manual(values=CONFIG$colors) +
    theme(axis.text.x=element_text(angle=90)) +
    facet_wrap(~chr)
```

## Total RPKM expression by chromosome (extracellular vs. intracellular)

```{r figure_x_total_chr_rpkm_expr_stage_specific}
ggplot(chr_expr_averages, aes(x=stage, y=total_expr, fill=stage)) + 
    geom_bar(stat='identity', color=CONFIG$stroke) +
    scale_fill_manual(values=CONFIG$colors) +
    theme(axis.text.x=element_text(angle=90)) +
    facet_wrap(~chr)
```

## Average RPKM expression by chromosome (developmental stage)

```{r figure_x_average_chr_gene_rpkm_expr_condition_specific}
# Compute average total RPKM for each chromosome (developmental stage)
chr_expr_averages <- rpkm_counts_long %>% 
    group_by(chr, condition) %>%
    summarize(average_expr=median(expr),
              total_expr=sum(expr))

ggplot(chr_expr_averages, aes(x=condition, y=total_expr, fill=condition)) + 
    geom_bar(stat='identity', color=CONFIG$stroke) +
    theme(axis.text.x=element_text(angle=90)) +
    facet_wrap(~chr)
```

## Total RPKM expression by chromosome (developmental stage)

```{r figure_x_total_chr_rpkm_expr_condition_specific}
ggplot(chr_expr_averages, aes(x=condition, y=total_expr, fill=condition)) + 
    geom_bar(stat='identity', color=CONFIG$stroke) +
    theme(axis.text.x=element_text(angle=90)) +
    facet_wrap(~chr)
```

## Average RPKM expression per nucleotide by PTU

```{r figure_x_average_rpkm_expr_per_nt_by_ptu}
# Add PTU mapping to RPKM gene counts
rpkm_totals <- merge(rpkm_totals, ptu_mapping, by.x='gene', by.y='gene_id')

# Determine the average expression by PTU
ptu_expr_averages <- rpkm_totals %>% 
    group_by(ptu) %>%
    summarise(average_expr=median(expr),
              total_expr=sum(expr))

# Add chromosome information
ptu_expr_averages <- merge(ptu_expr_averages, 
                           rpkm_totals %>% select(chr, ptu) %>% unique, by='ptu')

# Compute the combined length of CDS's for each PTU
ptu_cds_total_lengths <- merge(gene_lengths, ptu_mapping, by.x='gene', by.y='gene_id') %>%
    group_by(ptu) %>%
    summarise(total_length=sum(length))

ptu_stats <- merge(ptu_cds_total_lengths, ptu_expr_averages, by='ptu') %>%
    mutate(expr_per_nt=total_expr/total_length)

# ptu display order
ptu_order <- sort(unique(ptu_stats$ptu))

ggplot(ptu_expr_averages, aes(x=ptu, y=average_expr)) + 
    geom_bar(aes(fill=chr), stat='identity') +
    scale_x_discrete(limits=ptu_order) +
    theme(axis.text.x=element_blank(),
          axis.ticks.x=element_blank()) +
    xlab("PTU") +
    ylab("Median Gene Expression (RPKM)") +
    ggtitle("Average RPKM expression by PTU")
```

## Total RPKM expression per nucleotide by PTU

```{r figure_x_total_rpkm_expr_per_nt_by_ptu}
#ptu_stats <- merge(ptu_stats, rpkm_totals %>% select(chr, ptu) %>% unique, by='ptu')

# in order to avoid noise due to small, possibly incorrect PTU's, we will 
# exclude any PTUs with less than 5 genes
ptu_num_genes <- (ptu_mapping  %>% group_by(ptu) %>% summarize(num_genes=n()))
ptu_stats <- merge(ptu_stats, ptu_num_genes, by='ptu') %>%
    filter(num_genes >= 5)

ggplot(ptu_stats, aes(x=ptu, y=expr_per_nt)) +
    geom_bar(aes(fill=chr), stat='identity') +
    scale_x_discrete(limits=ptu_order) +
    theme(axis.text.x=element_blank(),
          axis.ticks.x=element_blank()) +
    xlab("polycistronic transcriptional unit (PTU)") +
    ylab("RPKM Expression (per nt)") +
    ggtitle("Average RPKM expression by PTU, scaled by PTU length")

```
