## Differences in trans-splicing acceptor site usage across developmental stage

For each pair of parasite developmental stages, a scatterplot showing the 
locations of primary SL sites which differ across stage is shown.

Further, genes which meet the following requirements and are thus are likely
candidates for SL switching are listed below:

1. Primary to secondary site coverage ratio > 2 for all stages
2. Minimum of 10 reads for all stages
3. Primary site location differs between at least two stages, and is located
   at least 20nt apart.

```{r figure_x_stage-specific_5utr_lengths, results='asis', fig.cap="**Figure SX** 5' UTR length across developmental stage.", warning=FALSE}
# create vectors to keep track of genes meeting each of the thresholds for at
# least one pair of stages compared.
hits_level1 <- c()
hits_level2 <- c()
hits_level3 <- c()
hits_level4 <- c()

for (s1_name in stages) {
    for (s2_name in stages) {
        # skip combined site dataframes and self comparisons
        if (s1_name == 'combined' || s2_name == 'combined' || s1_name == s2_name) {
            next
        }
        
        # create cross-stage dataframe and plot UTRs with variables lengths
        # across developmental stage
        dat <- create_utr_comparison_df(sl_sites, s1_name, s2_name)

        hits_level1 <- unique(c(hits_level1, dat$gene))

        # filter out genes which don't meet minimum criteria
        dat <- dat %>% filter(min_num_reads >= CONFIG$min_read_support_5utr)
        hits_level2 <- unique(c(hits_level2, dat$gene))

        cat('\n')

        print(plot_diff_utrs(dat, 'SL'))

        cat('\n')

        ## criteria
        dat <- dat %>% filter(s1_ptos > 2 & s2_ptos > 2)
        hits_level3 <- unique(c(hits_level3, dat$gene))

        # genes with stage-specific site switching
        dat <- dat %>% filter(len_diff > 20) 
        hits_level4 <- unique(c(hits_level4, dat$gene))
       
        # print list of genes
        dat %>%
            arrange(desc(min_num_reads)) %>%
            kable %>%
            print
    }
}

cat(sprintf('\n- Number genes with site switching across stages: %d\n', length(hits_level1)))
cat(sprintf('- Number with 3+ reads supporting each stage: %d\n', length(hits_level2)))
cat(sprintf('- Number with primary to secondary ratio > 2 for both stages: %d\n', length(hits_level3)))
cat(sprintf('- Number where acceptor sites used are >20nt apart: %d\n', length(hits_level4)))
```

## Differences in polyadenylation site usage across developmental stage

```{r figure_x_stage-specific_3utr_lengths, results='asis', fig.cap="**Figure SX** 3′ UTR length across developmental stage.", warning=FALSE}
for (s1_name in stages) {
    for (s2_name in stages) {
        # skip combined site dataframes and self comparisons
        if (s1_name == 'combined' || s2_name == 'combined' || s1_name == s2_name) {
            next
        }
        
        # create cross-stage dataframe and plot UTRs with variables lengths
        # across developmental stage
        dat <- create_utr_comparison_df(polya_sites, s1_name, s2_name)
        dat <- dat %>% filter(min_num_reads >= CONFIG$min_read_support_3utr)

        cat('\n')

        print(plot_diff_utrs(dat, 'Poly(A)'))

        cat('\n')

        # genes with stage-specific site switching
        dat %>% filter(s1_ptos > 2 & s2_ptos > 2 & len_diff >= 20 & min_num_reads >= 10) %>%
            arrange(desc(min_num_reads)) %>%
            kable %>%
            print
    }
}
```
