## Differences in trans-splicing acceptor site usage across developmental stage

For each pair of parasite developmental stages, a scatterplot showing the 
locations of primary SL sites which differ across stage is shown.

Further, genes which meet the following requirements and are thus are likely
candidates for SL switching are listed below:

1. Primary to secondary site coverage ratio > 2 for all stages
2. Minimum of 10 reads for all stages
3. Primary site location differs between at least two stages, and is located
   at least 20nt apart.

```{r figure_3x_stage-specific_5utr_lengths, results='asis', fig.cap="**Figure 3** 5' UTR length across developmental stage.", warning=FALSE}
# create vectors to keep track of genes meeting each of the thresholds for at
# least one pair of stages compared.
hits_level1 <- c()
hits_level2 <- c()
hits_level3 <- c()
hits_level4 <- c()

sl_plots <- list()

# list to store pairwise comparison tables
table_S8 <- list()

i <- 1

for (s1_name in stages) {
    for (s2_name in stages) {
        # skip combined site dataframes and self comparisons
        comparison <- paste(sort(c(s1_name, s2_name)), collapse='_vs_')

        # skip combined site dataframes, self comparisons, and previously
        # plotted pairs of stages
        if (s1_name == 'combined' || s2_name == 'combined' || 
            s1_name == s2_name || comparison %in% names(sl_plots)) {
            next
        }

        # create cross-stage dataframe and plot UTRs with variables lengths
        # across developmental stage
        dat <- create_utr_comparison_df(sl_sites, s1_name, s2_name)

        # Store unfiltered table results for output
        table_S8[[comparison]] <- dat

        # don't show genes with site-switches < 5nt
        dat <- dat %>% filter(len_diff > 5)

        hits_level1 <- unique(c(hits_level1, dat$gene))

        # filter out genes which don't meet minimum criteria
        dat <- dat %>% filter(min_num_reads >= CONFIG$min_read_support_5utr)
        hits_level2 <- unique(c(hits_level2, dat$gene))

        # plot utr lengths across stages
        sl_plots[[comparison]] <- plot_diff_utrs(dat, 'SL')

        ## criteria
        dat <- dat %>% filter(s1_ptos >= 3 & s2_ptos >= 3)
        hits_level3 <- unique(c(hits_level3, dat$gene))

        # genes with stage-specific site switching
        dat <- dat %>% filter(len_diff > 20) 
        hits_level4 <- unique(c(hits_level4, dat$gene))

        # add gene description and print list of interesting hits
        dat <- merge(dat, gene_info, by='gene')

        # display table of interesting hits
        caption <- sprintf("**Table S8%s** List of potentially interesting trans-splicing acceptor site switches comparing %s and %s.", 
                           letters[i], s1_name, s2_name)

        dat %>%
            filter(min_num_reads >= 10) %>%
            arrange(desc(min_num_reads)) %>%
            xkable(caption=caption) %>%
            print

        i <- i + 1
    }
}

cat(sprintf('\n- Number genes with site switching across stages (>5nt apart): %d\n', length(hits_level1)))
cat(sprintf('- Number with 3+ reads supporting each stage: %d\n', length(hits_level2)))
cat(sprintf('- Number with primary to secondary ratio > 2 for both stages: %d\n', length(hits_level3)))
cat(sprintf('- Number where acceptor sites used are >20nt apart: %d\n', length(hits_level4)))
```

## Differences in polyadenylation site usage across developmental stage

```{r figure_X4_stage-specific_3utr_lengths, results='asis', fig.cap="**Figure X4** 3′ UTR length across developmental stage.", warning=FALSE}
polya_plots <- list()

# list to store pairwise comparison tables
table_S9 <- list()

for (s1_name in stages) {
    for (s2_name in stages) {
        # skip combined site dataframes and self comparisons
        comparison <- paste(sort(c(s1_name, s2_name)), collapse='_vs_')

        # skip combined site dataframes, self comparisons, and previously
        # plotted pairs of stages
        if (s1_name == 'combined' || s2_name == 'combined' || 
            s1_name == s2_name || comparison %in% names(polya_plots)) {
            next
        }

        # create cross-stage dataframe and plot UTRs with variables lengths
        # across developmental stage
        dat <- create_utr_comparison_df(polya_sites, s1_name, s2_name)

        # Store unfiltered table results for output
        table_S9[[comparison]] <- dat

        # Filter out sites with low read support
        dat <- dat %>% filter(min_num_reads >= CONFIG$min_read_support_3utr)

        polya_plots[[comparison]] <- plot_diff_utrs(dat, 'Poly(A)')

        ## criteria
        dat <- dat %>% filter(s1_ptos >= 3 & s2_ptos >= 3)
        hits_level3 <- unique(c(hits_level3, dat$gene))

        # add gene description and print list of interesting hits
        dat <- merge(dat, gene_info, by='gene')

        # display table of interesting hits
        caption <- sprintf("**Table S9%s** List of potentially interesting polyadenylation site switches comparing %s and %s.", 
                           letters[i], s1_name, s2_name)

        dat %>%
            filter(len_diff >= 20 & min_num_reads >= 10) %>%
            arrange(desc(min_num_reads)) %>%
            xkable(caption=caption) %>%
            print

        cat('\n')
    }
}
```

## Stage-specific SL and Poly(A) site switch plots

Below, scatterplots of alternative trans-splicing and polyadenylation events
are shown for each pair of developmental stages.

```{r figure_3, warning=FALSE, fig.width=2160/192, fig.height=1080/192, dpi=192}
for (comparison in names(sl_plots)) {
    grid.arrange(sl_plots[[comparison]], polya_plots[[comparison]], ncol=2)
}
```
