## Differences in trans-splicing acceptor site usage across developmental stage

For each pair of parasite developmental stages, a scatterplot showing the 
locations of primary SL sites which differ across stage is shown.

Further, genes which meet the following requirements and are thus are likely
candidates for SL switching are listed below:

1. Primary to secondary site coverage ratio > 2 for all stages
2. Minimum of 20 reads for all stages
3. Primary site location differs between at least two stages, and is located
   at least 20nt apart.

```{r figure_x_stage-specific_5utr_lengths, results='asis', fig.cap="**Figure SX** 5' UTR length across developmental stage.", warning=FALSE}
for (s1_name in stages) {
    for (s2_name in stages) {
        # skip combined site dataframes and self comparisons
        if (s1_name == 'combined' || s2_name == 'combined' || s1_name == s2_name) {
            next
        }
        
        # create cross-stage dataframe and plot UTRs with variables lengths
        # across developmental stage
        dat <- create_utr_comparison_df(sl_sites, s1_name, s2_name, CONFIG$min_read_support_5utr)

        cat('\n')

        print(plot_diff_utrs(dat, 'SL'))

        cat('\n')

        # genes with stage-specific site switching
        dat %>% filter(s1_ptos > 2 & s2_ptos > 2 & len_diff >= 20 & min_num_reads > 20) %>%
            arrange(desc(min_num_reads)) %>%
            kable %>%
            print
    }
}
```

## Differences in polyadenylation site usage across developmental stage

```{r figure_x_stage-specific_3utr_lengths, results='asis', fig.cap="**Figure SX** 3′ UTR length across developmental stage.", warning=FALSE}
for (s1_name in stages) {
    for (s2_name in stages) {
        # skip combined site dataframes and self comparisons
        if (s1_name == 'combined' || s2_name == 'combined' || s1_name == s2_name) {
            next
        }
        
        # create cross-stage dataframe and plot UTRs with variables lengths
        # across developmental stage
        dat <- create_utr_comparison_df(polya_sites, s1_name, s2_name,
                                        CONFIG$min_read_support_3utr)

        cat('\n')

        print(plot_diff_utrs(dat, 'Poly(A)'))

        cat('\n')

        # genes with stage-specific site switching
        dat %>% filter(s1_ptos > 2 & s2_ptos > 2 & len_diff >= 20 & min_num_reads > 20) %>%
            arrange(desc(min_num_reads)) %>%
            kable %>%
            print
    }
}
```
